# (Markov Chain) Monte Carlo

## Monte Carlo Approximation

In Bayesian statistics, many inferential tasks involve 
computing some sort of expectation 
$$
\mathbb{E} h(Y)
=
\int h(y) p(y) dy.
$$

We may not know $\mathbb{E} h(Y)$ analytically, but as long as we 
can sample from $p(y)$ we should be able to (1) estimate $\mathbb{E} h(Y)$ well and (2) characterize the uncertainty in our estimate.

To do so we draw on our knowledge from Prob I.
Let $(Y_1, \dots, Y_m)$ be draws from $p(y).$ Then 
$$
\bar{h}_m
=
m^{-1} \sum\limits_{j=1}^{m} h(Y_j)
$$
has a number of nice properties which are implied by the strong law of large numbers and central limit theorem. In particular, provided that $\mathbb{E} \left[ h(Y)^2 \right] < \infty$, we have
$$
\begin{align*}
    \bar{h}_m &\stackrel{\textrm{a.s.}}{\rightarrow} \mathbb{E} h(Y) \\
    \mathbb{E}[\bar{h}_m] &= \mathbb{E} h(Y) \\
    \textrm{Var} \left( \bar{h}_m \right) &= m^{-1} \int_{\mathcal{Y}} ( h(y) - \mathbb{E}[h(Y)] )^2 f(y)dy \\
        &\approx m^{-1} \sum\limits_{j=1}^{m} ( h(Y_j) - \bar{h}_m )^2 := \hat{\sigma}^2_m \\
    \bar{h}_m &\rightarrow N \left( \mathbb{E} h(Y), m^{-1} \textrm{Var} [h(Y)] \right) \\
    &\approx N \left( \mathbb{E} h(Y), m^{-1} \hat{\sigma}^2_m \right)
\end{align*}
$$

So, $\bar{h}_m \pm 1.96 \frac{\hat{\sigma}_m}{\sqrt{m}}$ is 
an asymptotic 95% confidence interval for $\mathbb{E} h(Y)$. 

### Exercise #2: Monte Carlo approximation to the posterior predictive

Suppose we have a similar setup as Exercise #1, 
with $X_{1:n} \stackrel{\textrm{i.i.d.}}{\sim} \textrm{Exp}(\theta)$
and a prior of $\theta \sim \textrm{Gamma}(\alpha = 2, \beta = 2)$.

Now suppose we observe the following `x` vector:

```{r}
set.seed(1)
n <- 10
x <- runif(n)
x <- x / sum(x) * 10

x
```

**Part (a)**

Use a Monte Carlo algorithm to approximate the posterior predictive 
density $p(x_{n+1} \vert x_{1:n})$ at the point $x_{n+1} = 0$.

<details>
<summary style="font-weight: bold; cursor: pointer;">Solution</summary>

<br>

Recall from Exercise #1 that the Gamma is a conjugate prior for the Exponential$(\theta)$ 
data model, so we know that the posterior on $\theta$ is Gamma$(\alpha', \beta')$,
where $\alpha' = \alpha + n$ and $\beta' = \beta + \sum\limits_{i=1}^{n} x_i$.
From here, we can implement our Monte Carlo algorithm, repeatedly sampling $\theta$
from the Gamma$(\alpha', \beta')$ posterior and evaluating the exponential density at 0 for each posterior $\theta$ value.

```{r}
# set up parameters
n <- length(x)
alpha <- 2
beta <- 2

# use conjugacy (Exercise 1) to get alpha', beta' analytically
alpha_p <- alpha + n
beta_p <- beta + sum(x)

# repeatedly sample from posterior, then evaluate posterior predictive density at 0
n_draws <- 10^3
pp_draws <- rep(0, n_draws)
for (i in 1:n_draws) {
    # draw from posterior on theta
    posterior_theta <- rgamma(1, shape = alpha_p, rate = beta_p)

    # evaluate density at x = 0 given posterior theta
    pp_draws[i] <- dexp(0, rate = posterior_theta)
}

# plot histogram of posterior predictive
ggplot(data.frame(pp_draws), aes(x = pp_draws)) +
  geom_histogram(fill = "steelblue", color = "white", bins = 100) +
  labs(
    title = "Monte Carlo draws",
    x = "density",
    y = "frequency"
  ) +
  theme_minimal()

# get approximate posterior predictive at 0
a <- mean(pp_draws)
glue("Monte Carlo approximation: {a}")
```

</details>

**Part (b)**

Empirically verify the claim that $\bar{h}_m$ is asymptotically normal.
(A very rough visual verification will do).

<details>
<summary style="font-weight: bold; cursor: pointer;">Solution</summary>

<br>

We run the same procedure as in part (a), but over a large number of repeated simulations.

```{r}
n_sims <- 10^4
ests <- rep(0, n_sims)
for (j in 1:n_sims) {
    pp_draws <- rep(0, n_draws)
    for (i in 1:n_draws) {
        # draw from posterior on theta
        posterior_theta <- rgamma(1, shape = alpha_p, rate = beta_p)

        # evaluate density at x = 0 given posterior theta
        pp_draws[i] <- dexp(0, rate = posterior_theta)
    }
    ests[j] <- mean(pp_draws)
}
ggplot(data.frame(ests), aes(x = ests)) +
  geom_histogram(fill = "steelblue", color = "white", bins = 100) +
  labs(
    title = expression("Asymptotic distribution of" ~ bar(h)[m]),
    x = expression(bar(h)[m]),
    y = "frequency"
  ) +
  theme_minimal()
```

</details>

One can do better than naïve Monte Carlo sampling with more advanced approaches,
such as Importance Sampling.

Suppose, as above, that we have continuous random variables 
$(Y_1, \dots, Y_n)$ from a distribution with density $p(y)$.
Define a density $q$ on the same space where $p(y) > 0 \implies q(y) > 0$.

We don't assume that we know $p$ or $q$ directly, only that we know them each up to their own normalizing constant; we'll call the versions we know 
$\tilde{p}, \tilde{q}$ and define 
$$
\tilde{w}(y)
=
\begin{cases}
    \frac{\tilde{p}(y)}{\tilde{q}(y)}, &\text{ if } \tilde{q}(y) > 0 \\
    0 &\text{ if } \tilde{q}(y) = 0.
\end{cases}
$$

Then an importance sampling-based approximation to $\mathbb{E} h(Y)$ looks like
$$
\begin{align*}
    \mathbb{E} h(Y)
        &= \sum\limits_{i=1}^{n} \left( \frac{\tilde{w}(Z_i)}{ \sum\limits_{j=1}^{n} \tilde{w}(Z_j) } \right)
\end{align*},
$$
where $Z_{1:n} \stackrel{\textrm{i.i.d.}}{\sim} q$.
That is, we need to be able to sample from $q$, but we need only know the actual density up to a normalizing constant.

To minimize approximation error, we want to choose a $q$ such that $q(x) \approx h(x) p (x)$, though in general we err on the side of choosing a more diffuse (spread out) $q$.

## Markov Chain Monte Carlo

There are cases where we can't sample directly from $p$, and 
it seems difficult to choose a good $q$ -- this happens quite regularly in 
high-dimensional models.

The most common way to construct an MCMC algorithm is with something called the *Metropolis-Hastings algorithm*. Suppose we want to sample from $p(x)$ and for arbitrary $x'$ call $q(x \vert x')$ the proposal distribution (from which we can sample). For every $(x,x')$ pair we define the acceptance ratio
$$
\alpha(x', x)
=
\frac{p(x') q(x \vert x')}{p(x) q(x' \vert x)}.
$$

::: {.callout-note icon=false}
## Metropolis-Hastings

Initialize $x_0$ and draw $T$ samples as follows. 
For each $t \in \{1,\dots,T\}$:

**Step (1):** Sample $x \sim q(x \vert x_{t-1})$

**Step (2):** Sample $u \sim \textrm{Uniform}(0,1)$

**Step (3):** If $u < \alpha(x_{t-1}, x)$ then set $x_t = x$. Otherwise, set $x_t = x_{t-1}$
:::

MCMC is an incredibly general class of tools for sampling in these settings. A key difference between MC and MCMC algorithms is that the samples we get from MCMC are correlated. In practice, we often perform *thinning* (actually using only every $k^{th}$ MCMC sample) to reduce the correlation between our samples.

One of the most popular versions of Metropolis-Hasting is called the *Gibbs sampler*.
Suppose we have a $D$-dimensional parameter vector $\theta = (\theta_1, \dots, \theta_D)$ and want to sample from the posterior $p(\theta \vert x)$, but it's too complex to sample directly.

The Gibbs sampler iteratively samples from $p(\theta_d \vert x, \theta_{-d})$
for each $d \in \{1,\dots,D\}$, where $\theta_{-d}$ is the full $\theta$ 
without its $d^{th}$ component. As we iterate through the elements of $\theta$,
we condition on the updated values. That is, 

::: {.callout-note icon=false}
## Gibbs Sampling

Initialize $\theta^{(0)} \in \mathbb{R}^D$.
For each $t \in \{0,\dots,T-1\}$:

$$
\begin{align*}
    \theta_1^{(t+1)} 
        &:= \theta_1^* \sim p \left( \theta_1^{(t+1)} \vert x, \theta_{-1}^{(t)} \right) \\
    \theta_2^{(t+1)} 
        &:= \theta_2^* \sim p \left( \theta_2^{(t+1)} \vert x, \theta_1^{(t+1)}, \theta_{-(1:2)}^{(t)} \right) \\
        &\vdots \\
    \theta_d^{(t+1)}
        &:= \theta_d^* \sim p \left( \theta_d^{(t+1)} \vert x, \theta_{1:(d-1)}^{(t+1)}, \theta_{-(d+1:D)}^{(t)} \right) \\
        &\vdots \\
    \theta_D^{(t+1)}
        &=: \theta_D^* \sim p \left( \theta_D^{(t+1)} \vert x, \theta_{1:(D-1)}^{(t+1)} \right).
\end{align*}
$$
:::

### Exercise #3: Gibbs Sampling as Metropolis-Hastings

Each time we update a component of $\theta$, we are performing 
a Metropolis-Hastings move. Prove that this is the case.
It is sufficient to state, in the Gibbs sampling context, (i) what is $p$, (ii) what is $q$, and (iii) what is $\alpha$.

<details>
<summary style="font-weight: bold; cursor: pointer;">Solution</summary>

<br>

::: {.callout-note icon=false}
This solution comes almost verbatim from Gregory Gundersen's blog, which I highly recommend in general for anyone looking to learn more about fundamentals of Bayesian statistics. [@gundersen2020MH]
:::


In this case our target distribution $p$ is the full posterior 
$p(\theta \vert x)$ and, at any given step, our proposal $q$ is of the form $p \left( \theta_d^{(t+1)} \vert x, \theta_{1:(d-1)}^{(t+1)}, \theta_{-(d+1:D)}^{(t)} \right)$. 
When manipulating $p,q$ below, we will elide the superscript notation tracking the exact
iteration of the algorithm.

It will be important for us to note that 
$$
p(\theta \vert x)
=
p( \theta_d, \theta_{-d} \vert x )
=
p(\theta_d \vert x, \theta_{-d}) p(\theta_{-d} \vert x).
$$

In general, we know that 
$\alpha(\theta^*, \theta)=\frac{p(\theta) q(\theta^* \vert \theta)}{p(\theta^*) q(\theta \vert \theta^*)}$, where $p,q$ are the target and proposal described above. 
In our Gibbs sampling context, this becomes
$$
\begin{align*}
    \alpha(\theta^*, \theta)
        &=\frac{p(\theta^* \vert x) p(\theta_d \vert x,\theta_{-d})}
               {p(\theta \vert x) p(\theta^*_d \vert x,\theta^*_{-d})} \\
        &= \frac{
    \textcolor{blue}{p(\theta_d^\star \vert x, \theta^\star_{-d})}
    \textcolor{red}{p(\theta_{-d}^\star \vert x)}
    \textcolor{olive}{p(\theta_d \vert x, \theta_{-d})}
}{
    \textcolor{olive}{p(\theta_d \vert x, \theta_{-d})}
    \textcolor{red}{p(\theta_{-d} \vert x)}
    \textcolor{blue}{p(\theta_d^\star \vert x, \theta^\star_{-d})}
} \\
    &= 1.
\end{align*}
$$

So we have $\alpha(\theta^*, \theta) = 1$. In other words, Gibbs sampling uses 
Metropolis-Hastings moves where the acceptance probability is 1.

</details>




